---
title: "Lab1-UKN"
author: "U Kristine Nguyen"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

The working directory

```{r}
getwd()
```

## Task 2

Read in DDT data

```{r}
ddt <- read.csv("DDT.csv")
head(ddt)
```

## Task 3

  a. The qualitative variables are RIVER and SPECIES
  b. The quantitative variables are MILE, LENGTH, WEIGHT, and DDT

```{r}
sp <- ddt$SPECIES
tab <- table(sp)
barplot(tab, col = 1:3)
```

  c. There are 3 different SPECIES
  
  d. Subset the ddt data so that we only see observations with SPECIES, LMBASS and WEIGHT bigger than 800gms
```{r}
with(ddt, ddt[SPECIES == "LMBASS" & WEIGHT > 800,])
```
  
  e. Subset the ddt data so that we only see observations with RIVER, SCM and DDT bigger than 4.0
  
```{r}
with(ddt, ddt[RIVER == "SCM" & DDT > 4.0,])
```
  
## Clicker Questions

  1. Find the mean length of the fish in the ddt data frame
  
```{r}
lg <- ddt$LENGTH
mean(lg)
```
  
  2. Find the standard deviation of the weight of the fish in the ddt data frame
  
```{r}
wt <- ddt$WEIGHT
sd(wt)
```

  3. Yes
  
```{r}
plot(WEIGHT~LENGTH, data = ddt)
```

  4. If v=1:20, the last value of v/20 is 1.00

```{r}
v = 1:20
v/20
```

## Task 4

  a. Make a table of the rivers
  
```{r}
rr <- ddt$RIVER
table(rr)
```
  
  b. Make a color coded barplot of the rivers
  
```{r}
tabr <- table(rr)
barplot(tabr, col = 12:16)
```
  
  c. Make a table of the rivers crossed with the species of fish
  
```{r}
table(rr, sp)
```
  
  d. Make a barplot of the rivers crossed with the species of fish
  
```{r}
tabrs <- table(rr, sp)
barplot(tabrs, beside = T, col = rainbow(4))
```
  
## Task 5

  a. Make a pie chart of Fish SPECIES
  
```{r}
pie(tab)
```
  
  
  b. Make a pie chart of RIVERS
  
```{r}
pie(tabr)
```

## Task 6
  
  a. Make a boxplot of the variable DDT
  
```{r}
tdd <- ddt$DDT
boxplot(tdd, ylim = c(0, 1110))
```
  
  b. Make a boxplot of the variable WEIGHT
  
```{r}
boxplot(wt, ylim = c(170, 2310))
```
  
  c. Make a boxplot of the variable LENGTH
  
```{r}
boxplot(lg, ylim = c(15, 55))
```
  
## Task 7

  a. Make a coplot of LENGTH vs WEIGHT given RIVER
  
```{r}
rivcol=with(ddt, ifelse(RIVER == "FCM", "Red",
                        ifelse(RIVER == "LCM", "Orange",
                               ifelse(RIVER == "SCM", "Green", "Blue"))))
coplot(LENGTH~WEIGHT|RIVER, data = ddt, col = rivcol, ylim = c(0, 60))
```
  
  b. Make a coplot of DDT vs WEIGHT given SPECIES
  
```{r}
fishcol=with(ddt, ifelse(SPECIES == "CCATFISH", "Blue",
                        ifelse(SPECIES == "SMBUFFALO", "Purple", "Green")))
coplot(DDT~WEIGHT|SPECIES, data = ddt,col = fishcol, ylim = c(0, 65))
```
  
## Task 8

  a. Use ggplot and make the following boxplots
  
```{r}
library(ggplot2)
b = ggplot(ddt,aes(x = SPECIES, y = WEIGHT))
b = b + geom_boxplot(aes(fill = RIVER)) + labs(title = "U. Kristine Nguyen")
b
```
  
  b. Use ggplot and make the following violin plots
  
```{r}
v = ggplot(ddt, aes(x = RIVER, y = LENGTH))
v = v + geom_violin(aes(fill = SPECIES)) + labs(title = "U. Kristine Nguyen")
v
```
  
  c. Use ggplot and make the following scatter plots
  
```{r}
s = ggplot(ddt, aes(x = WEIGHT, y = LENGTH, fill = SPECIES))
s = s + geom_point(aes(colour = SPECIES)) + labs(title = "U. Kristine Nguyen")
s
```
