---
title: "Lab 2-ukn"
author: "U Kristine Nguyen"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

The working directory

```{r}
getwd()
```

## Task 2

Read in EPAGAS data

```{r}
epa <- read.csv("EPAGAS.csv")
head(epa)
```

## Task 3

  a. Verifying z-bar = 0 and (s_z)^2 = 1

```{r}
mpg <- epa$MPG
s <- sd(mpg)
m <- mean(mpg)
z = (mpg - m) / s
zbar <- mean(z)
sz2 = sum(((z-zbar)^2)/99)
zbar
sz2
```

  b. MPG possible outliers
  
```{r}
po <- mpg[abs(z) >= 2 & abs(z) <= 3]
po
```
  
  c. MPG outliers
  
```{r}
o <- mpg[abs(z) >= 3]
o
```
  
  d. Dotplot for outliers
  
```{r}
library(lattice)
ocol = with(epa, ifelse(mpg == po, "Blue",
                        ifelse(mpg == o, "Red", "Black")))
d <- data.frame(x = epa$MPG, y = c(1:100))
dotplot(y~x, data = d, col = ocol, xlab = "Miles per Gallon")
```

## Task 4

  a. Make a boxplot of mpg variable

```{r}
boxplot(mpg, horizontal = TRUE, notch = TRUE, col = "black", xlab = "Miles per Gallon", ylab = "Gas Mileage",title = "Mean Gas Mileage in Miles per Gallon")
```

  b. Using Chebyshev's theorem predict the proportion of data within 2 standard deviations of the mean of the mpg data
  
```{r}
ch = 1 - (1/4)
ch
```
  
  c. Calculate the exact proportion within 2 standard deviations of the mean
  
```{r}
sum(mpg > (m-(2*s)) & mpg < (m+(2*s))) / length(mpg)
```
  
  d. Does Chebyshev agree with the data?
  
    Chebyshev does not agree with the data.
  
  e. Using the empirical rule, what proportion of data should be within 2 standard deviations of the mean?
  
    Using the emperical rule, 0.95 of the data should be within 2 standard deviations of the mean.
  
  f. How well does it correspond?
  
    The exact proportion of the data within 2 standard deviations of the mean is 0.01 different from what is stated in the empirical rule.
  
  g. Is the empirical rule valid? Why?

    The empirical rule is valid because actual data with a pretty normal distribution fits the empirical rule pretty well. The experimental supports the theoretical.